The \emph{FrobeniusThresholds} package for \emph{Macaulay2}

### 2019-06-06 Fsing/FThresholdsPaper/FThresholds.tex
Updates to Section 2 of paper

{\color{red} This section will need to be modified so that it is consistent with the introduction section.  The point is that a lot of the notation repeated here was introduced earlier there.}

@@

As pointed out in the introduction, if $\mathfrak{m}$ is the homogeneous maximal ideal and $f \in \mathfrak{m}$, then for all $e$, the values $\nu^{\mathfrak{m}}_f(p^e)$ can be recovered from the $F$-pure threshold of $f$.  
This fact is used to speed up computations for certain polynomials whose $F$-pure thresholds can be computed quickly via specialized algorithms, namely diagonal polynomials, binomials, and forms in two variables (see Section \ref{sec.FPT}).  This feature can be disabled by setting the option {\tt UseSpecialAlgorithms} (default value {\tt true}) to {\tt false}.

%R = ZZ/17[x,y,z];
%f = x^3 + y^4 + z^5; -- a diagonal polynomial
%time nu(3, f)
%time nu(3, f, UseSpecialAlgorithms => false)

The function {\tt nu} works by searching through the list of potential integers $n$ and checking containments of the $n^\text{the}$ power of $I$ in a specified Frobenius power of $J$.  
The user can specify how this search is approached through the  option {\tt Search}.  Namely, a binary search is used when {\tt Search} is set to {\tt Binary} (and this is the default value), and a linear search is run when it is set to {\tt Linear}.

The option {\tt ReturnList} for {\tt nu} (whose default value is {\tt false}) can be set to {\tt true} so that instead of outputting $\nu_I^J(p^e)$, but a list contaning $\nu_I^J(p^i)$, for $i=0,\ldots,e$.
Alternatively, the option {\tt Verbose} can be set to  to request that the values $\nu_I^J(p^i)$ ($i=0,\ldots,e$) be printed as they are computed, to monitor the progress of the computation.

@@

Finally, suppose that {\tt ContainmentTest} is set to {\tt FrobeniusPower}.  Instead of producing the invariant $\nu_I^J(p^e)$, in this case the {\tt nu} function instead outputs the maximal integer $n$ such that the $n^{\textrm{th}}$ \emph{(generalized) Frobenius power} of $I$ is not contained in the $p^e$-th Frobenius power of $J$ \cite{hernandez+etal.frobenius_powers}. 
Note that the resulting value coincides with $\nu_I^J(p^e)$ when $I$ is replaced with a polynomial, or is a principal ideal, but need not coincide in general.
In line with the philosophy described in the introduction, after appropriately normalizing and taking the limit of these values as $e$ approaches infinity, these new values limit to certain invariants of interest called \emph{critical exponents} (see \emph{loc.~cit.}).

%R = ZZ/3[x,y];
%M = ideal(x, y);
%nu(3, M^5)
%nu(3, M^5, ContainmentTest => FrobeniusPower)

%R = ZZ/5[x,y,z];
%f = x^2*y^4 + y^2*z^7 + z^2*x^8;
%time nu(5, f) -- uses binary search (default)
%time nu(5, f, Search => Linear)
%M = ideal(x, y, z);
%time nu(2, M, M^2) -- uses binary search (default)
%time nu(2, M, M^2, Search => Linear) -- but linear seach gets luckier

%nu(5, f, Verbose => true)

>>>>

@@

As pointed out in the introduction, if $\mathfrak{m}$ is the homogeneous maximal ideal and $f \in \mathfrak{m}$, then for all $e$, the values $\nu^{\mathfrak{m}}_f(p^e)$ can be recovered from the $F$-pure threshold of $f$.  
This fact is used to speed up computations for certain polynomials whose $F$-pure thresholds can be computed quickly via specialized algorithms, namely diagonal polynomials, binomials, and forms in two variables (see Section \ref{sec.FPT}).  This feature can be disabled by setting the option {\tt UseSpecialAlgorithms} (default value {\tt true}) to {\tt false}.  Consider the following example, where $f$ is a diagonal polynomial.

i1 : R = ZZ/17[x,y,z];

i2 : f = x^3 + y^4 + z^5; -- a diagonal polynomial

i3 : time nu(3, f)
     -- used 0.00552312 seconds

o3 = 3756

o4 : QQ

i4 : time nu(3, f, UseSpecialAlgorithms => false)
     -- used 0.155826 seconds

o4 = 3756
\end{MyVerbatim}
}
\medspace

% R = ZZ/17[x,y,z];
% f = x^3 + y^4 + z^5; -- a diagonal polynomial
% time nu(3, f)
% time nu(3, f, UseSpecialAlgorithms => false)

We now describe the options available for the function {\tt nu}, and we focus primarily on the case where the second input is an ideal $I$, since the options work analogously when the second input is instead a polynomial $f$.

In general, the function {\tt nu} works by searching through the list of integers $n$, and checking containments of the $n^\text{th}$ power of $I$ in a specified Frobenius power of $J$.  
The user can specify how this search is approached through the  option {\tt Search}.  Namely, a binary search is used when {\tt Search} is set to {\tt Binary} (and this is the option's default value), and a linear search is run when it is set to {\tt Linear}.

i1 : R = ZZ/5[x,y,z];

i2 : f = x^2*y^4 + y^2*z^7 + z^2*x^8;

i3 : time nu(5, f) -- uses binary search (default)
     -- used 0.313778 seconds

o3 = 1124

i4 : time nu(5, f, Search => Linear)
     -- used 0.441903 seconds

o4 = 1124

i5 : m = ideal(x, y, z);

o5 : Ideal of R

i6 : time nu(2, m, m^2) -- uses binary search (default)
     -- used 1.58258 seconds

o6 = 97

i7 : time nu(2, m, m^2, Search => Linear) -- but linear seach gets luckier
     -- used 0.531175 seconds

o7 = 97
\end{MyVerbatim}
}
\medspace

%R = ZZ/5[x,y,z];
%f = x^2*y^4 + y^2*z^7 + z^2*x^8;
% time nu(5, f) -- uses binary search (default)
% time nu(5, f, Search => Linear)
% M = ideal(x, y, z);
% time nu(2, M, M^2) -- uses binary search (default)
% time nu(2, M, M^2, Search => Linear) -- but linear seach gets luckier

If the option {\tt ReturnList} is changed from its default value of {\tt false} to {\tt true}, instead of simply outputting $\nu_I^J(p^e)$, {\tt nu} outputs a list of the values $\nu_I^J(p^s)$, for $s=0,\ldots,e$.
Alternatively, the same information can be found by setting the option {\tt Verbose} to {\tt true} (its default value is {\tt false}), to request that each of these values $\nu_I^J(p^s)$ is printed as they are iteratively computed.  This can also serve a way to monitor the progress of the computation.

i8 : nu(5, f, ReturnList => true)

o8 = {0, 1, 8, 44, 224, 1124}

o8 : List

i9 : nu(5, f, Verbose => true)
nu(1) = 0
nu(p^1) = 1
nu(p^2) = 8
nu(p^3) = 44
nu(p^4) = 224
nu(p^5) = 1124
\end{MyVerbatim}
}

%nu(5, f, ReturnList => true)
%nu(5, f, Verbose => true)

@@

Finally, suppose that {\tt ContainmentTest} is set to {\tt FrobeniusPower}.  Instead of producing the invariant $\nu_I^J(p^e)$, in this case the {\tt nu} function instead outputs the maximal integer $n$ such that the $n^{\textrm{th}}$ \emph{(generalized) Frobenius power} of $I$ is not contained in the $p^e$-th Frobenius power of $J$ \cite{hernandez+etal.frobenius_powers}. 
Note that the resulting value coincides with $\nu_I^J(p^e)$ when $I$ is replaced with a polynomial, or is a principal ideal, but need not coincide in general. 
For example, the two commands below give different output values. 
{\small
\setstretch{.67}
\begin{MyVerbatim}

i1 : R = ZZ/3[x,y];

i2 : m = ideal(x, y);

i3 : nu(3, m^5)

o3 = 10

i4 : nu(3, m^5, ContainmentTest => FrobeniusPower)

o4 = 8
\end{MyVerbatim}
}
\medspace
\noindent In line with the philosophy described in the introduction, after appropriately normalizing and taking the limit of these values as $e$ approaches infinity, the values computed when {\tt ContainmentTest} is set to {\tt FrobeniusPower}
limit to certain invariants of interest called \emph{critical exponents} (see \cite{hernandez+etal.frobenius_powers}).

% R = ZZ/3[x,y];
% m = ideal(x, y);
% nu(3, m^5)
% nu(3, m^5, ContainmentTest => FrobeniusPower)
